Cultural Evolution of Conservation Behavior at the Intersection of Working

### 2020-12-10 DissertationProposal/DissertationProposal.Rmd
Test edits to intro and moon map

# Introduction
## Phenomenon
Human prosperity is unequivocally linked to the persistence of diverse, self-sustaining ecosystems through our dependence on ecosystem services and consumption of renewable resources. Until approximately 50 years ago, the rate of overall ecological regeneration exceeded our rates of consumption and destruction of earth’s renewable resources [@network2019global]. Today, on average, we degenerate these resources at a rate 56% faster than regeneration [@network2020global]. This is particularly concerning as renewable resources regenerate at a rate proportional to the resource stock remaining [@beltratti_sustainable_1993]. Thus, overuse of earth’s renewable resources creates a positive feedback loop where scarcity slows regeneration and further perpetuates scarcity [@regev_biological_1998].

The capacity of an ecosystem to provision renewable resources largely hinges on an area’s biodiversity [@haines2010links; @redford1995human]. As such, biodiversity is the key metric used to determine the integrity of an ecosystem and a key conservation target [@parrish_are_2003]. Each year, an ever growing proportion of earth’s marine and terrestrial areas are sectioned off to protect local biodiversity (hotspots), the flow of genetic diversity across a landscape (corridors), and human access to renewable resources (forest concessions, fisheries, etc.) (Figure \@ref(fig:paNumber)) [@iucn2020protectedplanet]. These area-based conservation initiatives are from hereon referred to as protected areas.

@@

Protected areas have proven to be a powerful tool for sustaining biodiversity and ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local communities [@adams_biodiversity_2004; @leverington_global_2010]. While high-profile protected area successes are salient in the conservation narrative, protected areas also commonly fail to reach their intended biodiversity targets. The ubiquity of these failures is somewhat speculative as unsuccessful protected areas go largely unreported [@bottrill_pursuit_2011; @sutherland_need_2004]. Regardless of the high degree of uncertainty in the prevalence of protected area failures, the cause is more often than not a lack of social buy-in and adherence to protected area regulations [@catalano_learning_2019].

Community norms around protected area adherence are characterized as individually costly, yet beneficial at the group level. Substantial variation exists in the degree to which communities adhere to protected area regulations; some communities develop strong and lasting conservation norms, while in others these norms degrade over time or may fail to emerge at all [@rogers_diffusion_2010; @simon_win-win_2012]. Given that the human dimensions are most often the crux of successful conservation, the field of conservation social science has been developed to better understand which social factors are pertinent to conservation success [@bennett_conservation_2017]. This body of work has shown that where group norms around the adoption of protected areas have emerged, they persist via group conformity, especially where clear group identities are present [@ostrom_governing_1990; @richerson2001evolution]. What is less understood however, is the emergence and initial adoption of conservation behaviors.

Research on the diffusion of conservation behaviors suffers from a lack of generalizability. Case studies tend to be highly context dependent and offer little mechanistic insight into the emergence and spread of these behaviors. Outside of conservation science, the field of evolutionary anthropology has devoted considerable effort to understanding how and why individuals adopt costly cooperative behaviors. Of particular relevance to understanding group adherence to protected area regulations, the theory of cultural group selection offers a mechanistic explanation for the emergence and spread of individually costly group norms.

## Primer on cultural group selection
The largely untested theory of cultural group selection poses a mechanism for the emergence and spread of individually costly yet group beneficial behaviors such as warfare or religious donations [@boyd_transmission_2010]. Cultural group selection suggests that groups who develop individually costly yet group beneficial norms will outcompete groups who do not. As an individual’s behavior is flexible, individuals may migrate from less to more successful groups or directly copy the behaviors they observe in visibly successful cooperative groups [@henrich_cultural_2004]. This change in the frequency of behavioral traits is analogous to the perpetuation of beneficial alleles in genetic natural selection. In these contexts, selective pressures still push individuals towards selfish behaviors like freeriding, but cooperative norms will emerge if group selection pressures outweigh those of individual selection overall [@mcelreath_mathematical_2008; @okasha_multilevel_2004; @price1972extension].

Existing cultural group selection theory is highly abstract, relying on simple, analytical models of nonspecific behaviors [@boyd_rapid_2011; @lehmann_cultural_2008]. Current theory therefore does not offer predictions about the emergence and adoption of conservation norms specifically. Conservation behaviors are fundamentally different from generic cooperative behaviors as they operate under the sink/source dynamics of renewable, mobile resources as well as variable resource regeneration rates. As such, cultural group selection theory must be extended to include these dynamics in order to be effectively applied to conservation.

## Applying cultural group selection to the adoption of conservation behaviors
Protected areas are an ideal microcosm for studying and applying cultural group selection. Protected areas have the potential to support resource sustainability and local livelihoods at the group level, but require costly cooperation from individuals [@adams_biodiversity_2004]. While this tradeoff is straightforward in theory, the social and ecological outcomes of protected area establishment vary significantly in practice [@naughton-treves_role_2005]. Agrawal and Redford [-@agrawal2006poverty] argue that this is because protected areas are often created as “a shot in the dark,” with little attention paid to the mechanisms which cause protected areas to be widely adopted and subsequently successful. Under cultural group selection theory, we expect that the key mechanism which drives the adoption of protected areas is the balance between the individual and group level selection pressures for cooperation or defection. Therefore, not only do protected areas represent a system where cultural group selection theory can be studied empirically, but principles of cultural group selection may also provide a way forward for predicting and shaping the success of protected areas on the ground.

## Fidelity of behavioral transmission
The final component of behavioral adoption that is necessary to understand and shape effective conservation is the fidelity of behavioral transmission. In situations where behaviors do emerge and spread within populations, behaviors are transmitted along a chain of individuals, where they may be modified considerably to the point where the behavioral outcome is unrecognizable [@lewis_transmission_2012; @mcguigan_over-imitation_2011]. This phenomenon is of particular interest to conservation science where the outcomes of behavioral adoption are of greater interest than the behaviors themselves. Theoretical insights and controlled empirical research on transmission fidelity suggests that behavioral outcomes are more likely to be retained during transmission when the copying individual has greater access to the demonstrator [@king2007use; @muthukrishna_when_2016]. These insights are crucial for practitioners aiming to implement long-lasting, effective conservation projects, but must first be tested empirically in a conservation context where an individual’s access to a behavioral demonstration is realistic and outcomes of adopted behaviors have real impacts for individuals, communities, and the environment.

## Dissertation objectives
The overarching question this research seeks to answer is: What are the underlying social mechanisms which cause conservation behaviors to emerge and spread efficiently within populations?

* Objective 1: Create a dynamic, agent-based model which extends theories of cultural group selection into the contexts of conservation. Formally test the intuition that between group imitation can lead to the spread of community wide protected area adoption whereas solitary groups may fail to cooperate under identical environmental constraints such as resource degradation, regeneration, and mobility.

* Objective 2: Calculate theoretical adoption probabilities of protected areas known as Community Forest Management Agreements (CoFMAs) for each ward on Pemba Island, Tanzania. Probabilities are based on the salience of group cultural identities and the costs and benefits of CoFMAs to individuals and groups. Compare theoretical predictions to the presence or absence of CoFMAs in reality.

* Objective 3: Quantitatively examine how social and physical access to a behavioral demonstration during behavioral transmission affects both the copying fidelity as well as the performance of copied behavior. This project will make use of the ongoing spread of fuel efficient cookstove technology on Pemba Island, Tanzania.

## Study System
All empirical research for this dissertation will take place on Pemba Island, Tanzania. In 2014, Pemba Island had a reported population of 406,808 [@RGZ2014Zanzibar]. This population is divided into wards of 2,000 to 6,000 individuals. The majority of Pemban wards are described as forest-dependent. The Department of Forests and Non-Renewable Natural Resources allows wards to enter into Community Forest Management Agreements (CoFMAs), which among other things, establish forest protected areas and exclude outside individuals [@benjaminsen2014between]. As of 2018, 18 Pemban wards had adopted CoFMAs [@andrews_cultural_2018].

@@

We can extend this model to explore how the sink/source dynamics between protected areas and adjacent working landscapes may provision more renewable resources than working landscapes alone. In this formation, we still have one homogenous group of agents who harvest regenerating resources from a landscape. This landscape now has some specified percentage sectioned off and inaccessible to the agents — the protected area. At every time step agents will attempt to harvest resources from the working landscape up until the specified maximum harvest amount. The outcome of this harvest attempt for each agent is the result of one random draw from a Binomial distribution where the number of trials is equal to the number of resources available on the working landscape divided by the number of individuals (i.e. that individual’s share of the available resources) and the probability of success is equal to the percent of an agent’s harvest time spent in the working landscape; in this scenario we are forcing obligatory cooperation so that percentage is always 100%. This will not always be the case as we build complexity in the model.

After the agents harvest resources from the landscape, the resources in the working and protected areas will reproduce by the specified regeneration factor. If the number of resources in either area exceeds the allotted carrying capacity, the excess resources will either die (i.e. not be available for agents in the following time step) or move to the other area (protected or working) with some specified probability also applied to a Binomial distribution. Resources which move to the adjacent area but exceed the carrying capacity will also die. If we set the probability of resource mobility to 100%, we can now calculate the maximum sustainable harvest amount for each agent where the only modification to the equation is the addition of mobile resources from the protected area (${P_{mob}}$) up until a specified carrying capacity ($W_{cc}$).

@@

Note that the sustainable harvest value is now dependent on the percent of the resources in the working landscape vs the protected area. We can run the model for every combination of protected area sizes and maximum harvest values to make sure that the model outcomes match out analytical intuition (fig \@ref(fig:abm2)). In the figure below, we can see that up until the maximum sustainable harvest value for this scenario (16), agents will always reach the set maximum harvest value until the resources are no longer made available to them (i.e. all protected). Above the maximum sustainable harvest value, the size of the protected area becomes greatly important for determining average harvest payoff to agents. In this scenario where the resources regenerate at a relatively high rate (20%) and the probability of mobility is 100%, the sink/source dynamics allow agents to reach the greatest average harvest value when the protected area is large and therefore provisioning many resources to the ever depleting working landscape at every time step.

@@

The probability of resource mobility is operationalized in the model by using random draws from a Binomial distribution. For every resource unit regenerated over the carrying capacity of the protected area, we take one random binomial draw with the probability of success equal to the probability of resource mobility. With all other model specifications kept equal and a maximum harvest value set slightly above the maximum sustainable harvest value for a working landscape alone, we can explore the interaction between protected area size and resource mobility given obligatory cooperation.

@@

To apply this model to conservation adoption and different mechanisms of social learning, we can allow agents to alter their harvest strategies beyond obligatory cooperation around protected area regulations. During model initialization, we can set the percentage of agents who we will allow to spend a proportion of their harvest effort in the protected area (defecting). The proportion of effort an individual agent spends in the protected area is drawn from the beta distribution below (fig \@ref(fig:betastart)) to ensure that the majority of effort is still spent in the working landscape at initialization.

@@

Agents modify their behavior (the percentage of time they spend in each area) at each time step according to one of two learning strategies. If agents are following a success-biased learning strategy, they are paired with another agent at random at each time step and compare strategies. If one agent performs worse than another in the previous time step, the underperforming agent will adopt a strategy in-between (split the time difference in half) their previous strategy and the strategy of the agent who outperformed them. If agents are following a conformist-biased strategy, they are grouped with 5 other agents at random at each time step, and adopt the mean strategy of all agents in their group. The learning strategy of the overall population is currently set to either success-biased or conformist-biased learning for all agents, but in future versions of the model, it will be an adjustable scale which assigns either a conformist-biased or success-biased learning strategy to a percentage of the agents. The entire model process is visualized conceptually in the figure below and a table of all the model starting conditions, key variables, and outcomes can be found in Table \@ref(tab:tabAbmVars)).

@@

With all model parameters kept constant as in the output shown in (fig \@ref(fig:abm3)), we can run the model with just 5% of the agents assigned as having the possibility to defect at initialization and all agents modifying harvest strategies via success-biased learning. The average final time step outcomes from five runs of this model can be seen in (fig \@ref(fig:abm4)).

```{r abm4, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) across five runs of the model. Outcomes are the result of many combinations of resource mobilities and protected area sizes. Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources. The regeneration factor at each time step is 1.20 and the maximum harvest value is set to 25. Five percent of agents assigned to defect at initialization. Agents learn through success-biased learning.", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/abmp4.png")
```
Allowing agents to modify harvest strategies and harvest from the protected area greatly reduces the overall harvest outcomes. To better understand the underlying dynamics, we set the starting model parameters to the ideal conditions for maximum payoff as identified in (fig \@ref(fig:abm3)) and observe the model behavior at each time step.

@@

```{r tabAbmVars, echo=F, fig.align='center'}
abmVars <- data.frame(Parameters = c("Individuals","Total Carrying Capacity", "Percent Carrying Capacity", 
                                    "Percent Protected", "Starting Cooperators", "Time Steps", "Max Harvest",
                                     "Resource Regeneration",  "Resource Mobility","Groups","Learning Strategy", 
                                     "Percent Cooperation", "Average Payoff", "Protected Area Degradation",
                                     "Working Landscape Degradation"), 
                      Type = c("Starting Condition","Starting Condition","Starting Condition",
                               "Starting Condition","Starting Condition","Starting Condition",
                               "Starting Condition","Key Independent Variable","Key Independent Variable",
                               "Key Independent Variable","Key Independent Variable",
                               "Outcome","Outcome","Outcome","Outcome"), 
                      Description = linebreak(c("Number of agents in simulation",
                                      "Maximum number of resources allowed on landscape",
                                      "Percent of carrying capacity at initialization",
                                      "Percent of landscape in protected area",
                                      "Starting percent of agents fully cooperating",
                                      "Number of time steps simulation will run for",
                                      "Maximum number of resources agents will harvest during one time step",
                                      "The percent of available resources which will regenerate at each time step in both protected and working landscapes",
                                      "The probability that any available resource will move from the  protected to working landscape or vice versa",
                                      "Solitary or multiple competing groups",
                                      "Assigns a percent of the population to success-biased imitation and the remaining population to conformist-biased imitation",
                                      "Percent of total population harvest time spent in the working landscape","Mean harvest for all agents at each time step",
                                      "Working landscape percent of total carrying capacity",
                                      "Protected area percent of total carrying capacity")))
kbl(abmVars,format="latex", caption = "Description of all model variables and outcomes",escape=F) %>% 
  kable_classic_2(full_width=F)%>% row_spec(0,bold=TRUE)%>%
    column_spec(1, width = "8em")%>%
    column_spec(2, width = "8em")%>%
  column_spec(3, width = "20em")

@@

### Methodological Specifics
I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi wards on Pemba Island, Tanzania (Fig \@ref(fig:mapp)). With the assistance of a translator, I will survey all the individuals responsible for constructing each cookstove, measure a suite of physical characteristics of each cookstove (length, height, etc.), and measure the fuel efficiency.

>>>>

# Introduction
Human prosperity is unequivocally linked to the persistence of diverse, self-sustaining ecosystems through our dependence on ecosystem services and consumption of renewable resources. Until approximately 50 years ago, the rate of overall ecological regeneration exceeded the rate at which we consume and degrade our environment [@network2019global]. Today, on average, we degenerate our renewable resources at a rate 56% faster than regeneration [@network2020global]. This is particularly concerning as renewable resources regenerate at a rate proportional to the resource stock remaining [@beltratti_sustainable_1993]. Thus, overuse of earth’s renewable resources creates a positive feedback loop where scarcity slows regeneration and further perpetuates scarcity [@regev_biological_1998].

Slowing the rate of loss and reduction of earth’s natural areas is decisively, the first step to stemming the global decline in natural resources and ecosystem function [@unep2018ngs; @haines2010links]. Achieving this will require both the adoption of new, efficient technologies and the preservation of currently intact areas [@secretariat2009review]. Toward this aim, each year we design and distribute technologies that use fewer or alternative resources such as improved efficiency cookstoves and solar panels [@ryan2012spreading]. In conjunction, we section off an ever growing proportion of earth’s marine and terrestrial areas to protect key resources (hotspots), the flow of resources across landscapes (corridors), and human access to renewable resources (forest concessions, fisheries, etc.) (Figure \@ref(fig:paNumber)) [@iucn2020protectedplanet].

@@

Both of these efforts benefit from a great deal of top-down support. Governments demarcate parks and wilderness areas, Western nonprofit organizations drop thousands of cookstoves across Equatorial Africa, and the global carbon market funds much of these efforts in the name of climate change mitigation [@redmond2015global; @wara2007global]. While they may be well supported in the onset, many conservation initiatives are intended only to seed a specific conservation paradigm or technology, relying on subsequent widespread community adoption for long-term benefits [@mills2019conservation; @sunderlin2012redd]. From hereon, we refer to geographic areas managed for conservation by communities themselves as *community conservation areas*, regardless of whether the *origin* of an area is community-based or top-down.

Community conservation areas can be a powerful tool for sustaining ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local communities [@adams_biodiversity_2004; @leverington_global_2010]. The success of these areas is largely contingent on local buy-in and adherence to area regulations — conservation doesn’t work if no one does it [@catalano_learning_2019]. Substantial variation exists in the degree to which communities adhere to conservation area regulations; some communities develop strong and lasting conservation norms, while in others these norms degrade over time or may fail to emerge at all [@simon_win-win_2012]. The degree to which conservation areas fail to catch on in communities is likely much greater than we recognize; unsuccessful conservation projects go largely unreported [@bottrill_pursuit_2011; @sutherland_need_2004].

Similar to community-based conservation areas, new efficient technologies have the potential to mitigate natural resource degradation while benefiting human populations, but rely greatly on community buy-in. Adoption of these technologies may diffuse and become ubiquitous within a population [@bollinger2012peer]. This is, however, not always the case. The reasons for if and when an individual adopts a new technology, especially in the contexts of conservation, are often unclear [@smith2014early]. Even in cases where a technology does spread within a population, the specific use behavior and use outcome may become so greatly modified along the transmission chain that the population level result is unrecognizable from the original intent [@lewis_transmission_2012; @mcguigan_over-imitation_2011]. Theoretical insights and controlled empirical research give suggestions for how intended technological adoption outcomes may be more likely to be retained during transmission from individual to individual or group to group. So far, these insights are yet to be validated empirically in a conservation context where technology adoption has real impacts for individuals, communities, and the environment, and those impacts may affect future adoptions [@king2007use; @muthukrishna_when_2016].

The efficacy of both community conservation areas and improved technologies depends on widespread behavioral adoption at both the individual and community level. Cultural group selection theory poses mechanisms for the spread and persistence of behaviors within and between groups [@boyd_transmission_2010]. Under this theory, groups develop characteristics or institutions such as a resource management agreement or community conservation norms and those characteristics can spread between groups through one of three mechanisms. Groups may outcompete or kill other groups with different institutions, individuals in one group may emigrate to another group, or individuals in less successful groups may copy behaviors they observe in more successful groups [@henrich_cultural_2004; @richerson_cultural_2016]. This change in the frequency of behavioral traits in groups is analogous to the spread and perpetuation of beneficial alleles between mixing groups in genetic natural selection. In these contexts, selective pressures still act on individuals and push them to engage in non-cooperative behaviors like stealing resources from a community conservation area, but they also experience selective pressure at the group level which may lead to cooperative behavior even at some individual cost [@mcelreath_mathematical_2008; @okasha_multilevel_2004; @price1972extension].

Researchers have posited that cultural group selection theory should be applicable to understanding the spread and persistence of conservation behaviors such as adherence to community conservation areas or adoption of new efficient technologies [@kline2018designing]. Cultural group selection theory offers mechanistic insights for the emergence and spread of behaviors within and between groups broadly, but its applicability to conservation specifically is unknown. Conservation has many aspects which make it unique compared to behavioral adoption in the abstract. In conservation, an individual’s behavior directly affects their environment, which may in turn become altered in a way which shapes future individual decision making. Feedbacks between individual behavior and resource availability are further complicated by resource characteristics such as carrying capacity, mobility, and regeneration rate. These specifics make conservation a fundamentally different system from general behavioral adoption and cooperation, thus we must extend cultural group selection theory considering conservation explicitly. The aim of this research is to expand the theories of behavioral adoption via cultural group selection to explicitly consider factors unique to conservation, and to empirically examine the mechanisms of conservation behavioral adoption within and between groups. This research will offer insights for strategically seeding both community conservation areas and new technologies so that they may offer the greatest benefit to communities and the preservation of renewable resources.

## Dissertation objectives
The overarching question this research seeks to answer is: What are the underlying social mechanisms which cause conservation behaviors to emerge, spread, and persist within populations?

* Objective 1: Create an agent based model to theoretically test whether cultural group selection via success copying or direct competition results in greater conservation adoption and better environmental outcomes, particularly in environmental contexts where solitary groups fail to cooperate.

* Objective 2: Empirically test whether imitating the successes of adjacent groups or competing against them for resources is driving the spread of community conservation areas on Pemba Island, Tanzania.

* Objective 3: Examine if the efficiency of an improved cookstove design has been modified as it has spread through three communities on Pemba Island, Tanzania. Quantitatively assess the relationship between transmission fidelity and the social dynamics between demonstrating/copying individuals.

## Study System
All empirical research for this dissertation will take place on Pemba Island, Tanzania; Pemba is one of two main islands in the Zanzibar archipelago (Figure \@ref(fig:moon)). In 2014, Pemba Island had a reported population of 406,808 [@RGZ2014Zanzibar]. This population is divided into 121 wards of 2,000 to 6,000 individuals. The majority of Pemban wards are described as forest-dependent, with 90% of households relying exclusively on forest products (fuelwood and charcoal) to meet their daily cooking needs. The yearly deforestation rate across all of Zanzibar was 1.1% in 2014 and is worsening as the population increases by 3.2% each year [@RGZ2014Zanzibar]. On Pemba, the fuelwood offtake around the cities is approximately ten times that of sustainable harvest [@ely2000participatory]. Unlike many areas with high deforestation, on Pemba, 86% of all forest loss is attributed to the local communities directly, rather than outside entities (e.g. foregin corporations) [@terracarbonreport].

All of the above factors contributed to Zanzibar being selected as one of eight pilot sites in Tanzania for the REDD+ program in 2009 [@burgess2010getting]. As part of this program on Pemba, Zanzibar’s Department of Forests and Non-Renewable Natural Resources worked with 18 wards to establish Community Forest Management Agreements (CoFMAs) [@andrews_cultural_2018]. Crucially, these agreements establish both a *use area* and a *conservation area* inside the forest in that ward and are meant to exclude outside individuals from harvesting inside the forest as well (Figure \@ref(fig:moon)) [@benjaminsen2014between].

```{r moon, echo=FALSE, fig.cap="Pemba Island, Tanzania. Showing wards with and without CoFMAs. For each CoFMA, the size (ha) is proportional to the size of the circle and the percent area designated as Use and Conservation is indicated by the blue and green colors respectively.", out.width = "50%", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/MoonMapEditedNoLabs.png")
```

A secondary component of the REDD+ program on Pemba was the distribution of fuel efficient cookstoves in three wards [@terracarbonreport]. These cookstoves have been shown to reduce fuelwood use by up to 50% [@lung_large-scale_2019]. The specific type of stove that was distributed can be built with materials readily available on Pemba. The intent of this aspect of the REDD+ project was to seed this design so that it may be widely copied throughout the communities and reduce deforestation beyond that of the original cookstoves.

The final key component of the REDD+ project as it relates to this research was the promise of payments to communities for their conservation efforts. This aspect of the program never came to fruition due to issues with the carbon validation process [@andrewsdoes]. Given that this aspect was unrealized, the CoFMAs have no outside budget and are now entirely community run, although still designated by the Department of Forests and Non-Renewable Natural Resources. 
Since the establishment of the original 18 CoFMAs, ten more wards have elected to enter the process of establishing CoFMAs, even without the promise of community payments from REDD+ [@andrewsdoes]. Furthermore, the seeded cookstove design has continued to spread across communities. While continuing to spread, both the CoFMAs and cookstoves have had variable success in reducing deforestation and fuelwood use respectively. This variation in both spread and efficacy of both of these initiatives within and between communities serves as the basis for the empirical research in this dissertation.

@@

We can extend this model to explore how the sink/source dynamics between protected areas and adjacent working landscapes may provision more renewable resources than working landscapes alone. In this formation, we still have one homogenous group of agents who harvest regenerating resources from a landscape. This landscape now has some specified percentage sectioned off and inaccessible to the agents — the protected area. At every time step agents will attempt to harvest resources from the working landscape up until the specified maximum harvest amount. The outcome of this harvest attempt for each agent is the result of one random draw from a binomial distribution where the number of trials is equal to the number of resources available on the working landscape divided by the number of individuals (i.e. that individual’s share of the available resources) and the probability of success is equal to the percent of an agent’s harvest time spent in the working landscape; in this scenario we are forcing obligatory cooperation so that percentage is always 100%. This will not always be the case as we build complexity in the model.

After the agents harvest resources from the landscape, the resources in the working and protected areas will reproduce by the specified regeneration factor. If the number of resources in either area exceeds the allotted carrying capacity, the excess resources will either die (i.e. not be available for agents in the following time step) or move to the other area (protected or working) with some specified probability also applied to a binomial distribution. Resources which move to the adjacent area but exceed the carrying capacity will also die. If we set the probability of resource mobility to 100%, we can now calculate the maximum sustainable harvest amount for each agent where the only modification to the equation is the addition of mobile resources from the protected area (${P_{mob}}$) up until a specified carrying capacity ($W_{cc}$).

@@

Note that the sustainable harvest value is now dependent on the percent of the resources in the working landscape vs the protected area. We can run the model for every combination of protected area sizes and maximum harvest values to make sure that the model outcomes match out analytical intuition (figure \@ref(fig:abm2)). In the figure below, we can see that up until the maximum sustainable harvest value for this scenario (16), agents will always reach the set maximum harvest value until the resources are no longer made available to them (i.e. all protected). Above the maximum sustainable harvest value, the size of the protected area becomes greatly important for determining average harvest payoff to agents. In this scenario where the resources regenerate at a relatively high rate (20%) and the probability of mobility is 100%, the sink/source dynamics allow agents to reach the greatest average harvest value when the protected area is large and therefore provisioning many resources to the ever depleting working landscape at every time step.

@@

The probability of resource mobility is operationalized in the model by using random draws from a binomial distribution. For every resource unit regenerated over the carrying capacity of the protected area, we take one random binomial draw with the probability of success equal to the probability of resource mobility. With all other model specifications kept equal and a maximum harvest value set slightly above the maximum sustainable harvest value for a working landscape alone, we can explore the interaction between protected area size and resource mobility given obligatory cooperation.

@@

To apply this model to conservation adoption and different mechanisms of social learning, we can allow agents to alter their harvest strategies beyond obligatory cooperation around protected area regulations. During model initialization, we can set the percentage of agents who we will allow to spend a proportion of their harvest effort in the protected area (defecting). The proportion of effort an individual agent spends in the protected area is drawn from the beta distribution below (figure \@ref(fig:betastart)) to ensure that the majority of effort is still spent in the working landscape at initialization.

@@

Agents modify their behavior (the percentage of time they spend in each area) at each time step according to one of two learning strategies. If agents are following a success-biased learning strategy, they are paired with another agent at random at each time step and compare strategies. If one agent performs worse than another in the previous time step, the underperforming agent will adopt a strategy in-between (split the time difference in half) their previous strategy and the strategy of the agent who outperformed them. If agents are following a conformist-biased strategy, they are grouped with 5 other agents at random at each time step, and adopt the mean strategy of all agents in their group. The learning strategy of the overall population is currently set to either success-biased or conformist-biased learning for all agents, but in future versions of the model, it will be an adjustable scale which assigns either a conformist-biased or success-biased learning strategy to a percentage of the agents. The entire model process is visualized conceptually in the figure below and a table of all the model starting conditions, key variables, and outcomes can be found in Table \@ref(tab:tabAbmVars).

@@

With all model parameters kept constant as in the output shown in (figure \@ref(fig:abm3)), we can run the model with just 5% of the agents assigned as having the possibility to defect at initialization and all agents modifying harvest strategies via success-biased learning. The average final time step outcomes from five runs of this model can be seen in (figure \@ref(fig:abm4)).

```{r abm4, out.width = "75%",echo=FALSE, fig.cap="Average harvest values for all agents in the final time step (30) across five runs of the model. Outcomes are the result of many combinations of resource mobilities and protected area sizes. Model starting conditions include 100 individuals harvesting from a landscape of 10000 resources. The regeneration factor at each time step is 1.20 and the maximum harvest value is set to 25. Five percent of agents assigned to defect at initialization. Agents learn through success-biased learning.", fig.align='center', dpi=72, fig.pos= "H"}
knitr::include_graphics("Figures/Ch1/abmp4.png")
```
Allowing agents to modify harvest strategies and harvest from the protected area greatly reduces the overall harvest outcomes. To better understand the underlying dynamics, we set the starting model parameters to the ideal conditions for maximum payoff as identified in (figure \@ref(fig:abm3)) and observe the model behavior at each time step.

@@

```{r tabAbmVars, echo=F, fig.align='center'}
abmVars <- data.frame(Parameters = c("Individuals","Total Carrying Capacity", "Percent Carrying Capacity", 
                                    "Percent Protected", "Starting Cooperators", "Time Steps", "Max Harvest",
                                     "Resource Regeneration",  "Resource Mobility","Groups","Learning Strategy", 
                                     "Percent Cooperation", "Average Payoff", "Protected Area Degradation",
                                     "Working Landscape Degradation"), 
                      Type = c("Starting Condition","Starting Condition","Starting Condition",
                               "Starting Condition","Starting Condition","Starting Condition",
                               "Starting Condition","Key Independent Variable","Key Independent Variable",
                               "Key Independent Variable","Key Independent Variable",
                               "Outcome","Outcome","Outcome","Outcome"), 
                      Description = linebreak(c("Number of agents in simulation",
                                      "Maximum number of resources allowed on landscape",
                                      "Percent of carrying capacity at initialization",
                                      "Percent of landscape in protected area",
                                      "Starting percent of agents fully cooperating",
                                      "Number of time steps simulation will run for",
                                      "Maximum number of resources agents will harvest during one time step",
                                      "The percent of available resources which will regenerate at each time step in both protected and working landscapes",
                                      "The probability that any available resource will move from the  protected to working landscape or vice versa",
                                      "Solitary or multiple competing groups",
                                      "Assigns a percent of the population to success-biased imitation and the remaining population to conformist-biased imitation",
                                      "Percent of total population harvest time spent in the working landscape","Mean harvest for all agents at each time step",
                                      "Working landscape percent of total carrying capacity",
                                      "Protected area percent of total carrying capacity")))
kbl(abmVars,format="latex", caption = "Description of all model variables and outcomes",escape=F,booktabs = T) %>% 
  kable_classic_2(full_width=F)%>% row_spec(0,bold=TRUE)%>%
    column_spec(1, width = "8em")%>%
    column_spec(2, width = "8em")%>%
  column_spec(3, width = "20em")%>%kableExtra::kable_styling(latex_options = "striped")

@@

### Methodological Specifics
I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi wards on Pemba Island, Tanzania (figure \@ref(fig:mapp)). With the assistance of a translator, I will survey all the individuals responsible for constructing each cookstove, measure a suite of physical characteristics of each cookstove (length, height, etc.), and measure the fuel efficiency.
